Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2020-04-15 thirdPaper/draft/typeConeAssociahedra.tex
minor changes type cones

% Type cone
\newcommand{\typeCone}{\mathbb{TC}} % type cone
\newcommand{\ctypeCone}{\overline{\mathbb{TC}}} % type cone
\newcommand{\compatibilityDegree}[2]{(#1\,\|\,#2)} % compatibility degree
\newcommandx{\coefficient}[3][1={\ray[s]}, 2=\ray, 3=\ray']{\alpha_{#2,#3}(#1)} % coefficient in linear dependence

@@

Moreover, we use a similar approach to study the space of polytopal realizations of the $\b{g}$-vector fans of another generalization of the associahedron: non-kissing complexes (a.k.a. support $\tau$-tilting complexes) of gentle algebras. We show that the space of realizations of the non-kissing fan is simplicial when the gentle bound quiver is brick and $2$-acyclic, and we describe in this case its facet-defining inequalities in terms of mesh mutations.

@@

In the present paper, we explore specifically \textbf{Gentle associahedra}. These complexes where constructed in~\cite{PaluPilaudPlamondon-nonkissing} as polytopal realizations of finite non-kissing complexes of gentle bound quivers. They encompass two families of simplicial complexes studied independently in the literature: on the one hand the grid associahedra introduced by T.~K.~Petersen, P.~Pylyavskyy and D.~Speyer in~\cite{PetersenPylyavskyySpeyer} for a staircase shape, studied by F.~Santos, C.~Stump and V.~Welker~\cite{SantosStumpWelker} for rectangular shapes, and extended by T.~McConville in~\cite{McConville} for arbitrary grid shapes; and on the other hand the Stokes polytopes and accordion associahedra studied by Y.~Baryshnikov~\cite{Baryshnikov}, F.~Chapoton~\cite{Chapoton-quadrangulations}, A.~Garver and T.~McConville~\cite{GarverMcConville} and T.~Manneville and V.~Pilaud~\cite{MannevillePilaud-accordion}. The latter, sometimes called accordiohedra, have also~recently appeared in the mathematical physics litterature concerning scattering amplitudes~\cite{BanerjeeLaddhaRaman-StokesPolytopes, Raman-PositiveGeometry}, with realizations inspired by those in~\cite{ArkaniHamedBaiHeYan}. It was shown in~\cite{PaluPilaudPlamondon-nonkissing, BrustleDouvilleMousavandThomasYildirim} that the non-kissing complex of a gentle bound quiver provides a combinatorial model for the support $\tau$-tilting complex~\cite{AdachiIyamaReiten} of the associated gentle algebra. Even for finite non-kissing complexes, it turns out that the type cone of the non-kissing fan is not always simplicial (see \cref{exm:nonSimplicialTypeConeNKC}). We prove however that the type cone is always simplicial when the gentle bound quiver is brick (at least $2$ relations in any cycle) and $2$-acyclic (no oriented cycle of length~$2$). Moreover, we precisely describe the facet-defining inequalities of the type cone in terms of linear dependences of $\b{g}$-vectors involved in certain mesh relations. We thus automatically derive a construction similar to that of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} describing all polytopal realizations of the $\b{g}$-vector fans of these brick and $2$-acyclic gentle bound quivers as sections of a positive orthant by well-chosen affine spaces (see \cref{subsec:typeConeCA}). Again, this result mainly relies on finding which relations minimally generate the relations between the $\b{g}$-vectors in the non-kissing complex. We present both a purely combinatorial proof (see \cref{subsec:typeConeCA}) and a purely algebraic proof for (non-necessarily gentle) brick algebras making use of extriangulated categories (see \cref{sec:extricats}).

@@

\smallskip
Further families of generalizations of the associahedron certainly deserve to be studied with our type cone approach.
We are investigating in particular graph associahedra~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}, brick polytopes~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and quotientopes~\cite{Reading-latticeCongruences, Reading-CambrianLattices, PilaudSantos-quotientopes}.
In fact, as already mentioned, it seems sometimes easier to find all polytopal realizations of a fan (by describing its type cone in terms of certain specific pairs of adjacent maximal cones) than to identify one specific explicit realization.
This is clearly the case when the type cone is simplicial as illustrated by the results of this paper: we naturally obtained all polytopal realizations of the $\b{g}$-vector fans of cluster algebras from cyclic seeds and of gentle algebras whose polytopality was only established very recently~\cite{HohlwegPilaudStella, PaluPilaudPlamondon-nonkissing}.
% Even when the type cone is not simplicial, describing the type cone sometimes leads to relevant explicit realizations of the fan, as we illustrate on graph associahedra (see \cref{exm:constructionsGA}).
We believe that this approach gives reasonable hope that the polytopality of further fans could in the future be established using our type cone approach.
Tempting candidates are quotientopes for hyperplane arrangements beyond type~$A$, which are particularly interesting since the coarsenings of a fan essentially correspond to the faces of its type cone (see \cref{subsec:facesTypeCone}).

@@

\newtheorem*{coro:simplicialTypeCone}{\cref{coro:simplicialTypeCone}}
\begin{coro:simplicialTypeCone}
\label{coro:simplicialTypeCone}
Assume that the type cone~$\typeCone(\Fan)$ is simplicial and let~$\b{K}$ be the $(N-n) \times N$-matrix whose rows are the inner normal vectors of the facets of~$\typeCone(\Fan)$. Then the polytope
\[
R_\b{\ell} \eqdef \set{\b{z} \in \R^N}{K\b{z} = \b{\ell} \text{ and } \b{z} \ge 0}
\]
is a realization of the fan~$\Fan$ for any positive vector~$\b{\ell} \in \R_{>0}^{N-n}$.
Moreover, the polytopes~$R_\b{\ell}$ for~$\b{\ell} \in \R_{>0}^{N-n}$ describe all polytopal realizations of~$\Fan$.
\end{coro:simplicialTypeCone}

@@

\para{Schematic diagram of logical dependencies}

@@

\vspace{.5cm}
\tikzset{
  double arrow/.style={
    -latex, line width=3pt, black, % first arrow
    postaction={draw,line width=2pt, white, shorten <= 1,shorten >= 2}, % second arrow
  }
}
\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
	\node[diamond,aspect=3, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=-2] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center, minimum width = 5.3cm, minimum height=1.2cm] (I21) at (-6,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5.3cm, minimum width = 5cm, minimum height=1.2cm] (I22) at (0,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5.3cm, minimum width = 5cm, minimum height=1.2cm] (I23) at (6,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
% 	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse,
% 	rectangle,rounded corners=10, 
	draw=orange, fill=orange!10, align=center, text width=5cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse,
% 	rectangle,rounded corners=10,
	draw=green, fill=green!10, align=center, text width=5cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(-1.6,-1.5)}]\textit{applied to}}]{} (I21.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=3}, draw=blue, fill=blue!10] (I1.270) to node[label={[shift={(1,-1.)}]\textit{applied to}}]{} (I22.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.360) to [bend left, looseness=1.1] node[label={[shift={(1.6,-1.5)}]\textit{applied to}}]{} (I23.90);
% 	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
% 	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
% 	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{\textit{used for}} (I22.270)+(0,.15);
% 	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{\textit{used for}} (I23.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies, in=270] (I22.270)+(-1.5,-1.2) to node[semithick, right, outer sep=0, pos=0.55]{\textit{used for}} ([yshift=.2cm] I22.270);
	\draw[thick, double distance=2pt, -implies, in=270] (I23.270)+(-1.5,-1.2) to node[semithick, right, outer sep=0, pos=0.55]{\textit{used for}} ([yshift=.2cm] I23.270);
	\draw[thick, ->] (II1.0) to node[semithick, below, outer sep=0]{\textit{generalizes to}} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=([shift={(0,-.15)}]I22.270), \p{right}=(I23.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I23.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
}

%\centerline{
%\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
% 	\node[cloud, cloud puffs=25, cloud ignores aspect, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=3, outer sep=5] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
%	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
%	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
%	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
%	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
%	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
%	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
% 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
% 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.240) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
% 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.300) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
% 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
% 	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
%	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
%	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
%	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes to} (II2.180);
%	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=([shift={(0,-.15)}]I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
%	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
%	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
%	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
%\end{tikzpicture}
%}
%
%\medskip
%
%\begin{center}
%\includegraphics[scale=.9]{Schema}
%\end{center}

@@

A hyperplane~$H \subset \R^d$ is a \defn{supporting hyperplane} of a set~$X \subset \R^d$ if~$H \cap X \ne \varnothing$ and~$X$ is contained in one of the two closed half-spaces of~$\R^d$ defined by~$H$.

We denote by~$\R_{\ge0}\rays \eqdef \set{\sum_{\ray \in \rays} \lambda_{\ray} \, \ray}{\lambda_{\ray} \in \R_{\ge0}}$ the \defn{positive span} of a set~$\rays$ of vectors of~$\R^d$.
A \defn{polyhedral cone} is a subset of~$\R^d$ defined equivalently as the positive span of finitely many vectors or as the intersection of finitely many closed linear halfspaces.
The \defn{faces} of a cone~$C$ are the intersections of~$C$ with the supporting hyperplanes of~$C$.
The $1$-dimensional (resp.~codimension~$1$) faces of~$C$ are called~\defn{rays} (resp.~\defn{facets}) of~$C$.
A cone is \defn{simplicial} if it is generated by a set of linearly independent vectors.

A \defn{polyhedral fan} is a collection~$\Fan$ of polyhedral cones such that
\begin{itemize}
\item if~$C \in \Fan$ and~$F$ is a face of~$C$, then~$F \in \Fan$,
\item the intersection of any two cones of~$\Fan$ is a face of both.
\end{itemize}
A fan is \defn{simplicial} if all its cones are, \defn{complete} if the union of its cones covers the ambient space~$\R^d$, and \defn{essential} if it contains the cone~$\{\b{0}\}$. Note that every complete fan is the product of an essential fan with its lineality space.
For two fans~$\Fan, \cal{G}$ in~$\R^d$, we say that~$\Fan$ \defn{refines}~$\cal{G}$ (and that~$\cal{G}$ \defn{coarsens}~$\Fan$) if every cone of~$\Fan$ is contained in a cone of~$\cal{G}$.

A \defn{polytope} is a subset~$P$ of~$\R^d$ defined equivalently as the convex hull of finitely many points or as a bounded intersection of finitely many closed affine halfspaces.
The \defn{dimension}~$\dim(P)$ is the dimension of the affine hull of~$P$.
The \defn{faces} of~$P$ are the intersections of~$P$ with its supporting hyperplanes.
The dimension~$0$ (resp.~dimension~$1$, resp.~codimension~$1$) faces are called \defn{vertices} (resp.~\defn{edges}, resp.~\defn{facets}) of~$P$.
A polytope is \defn{simple} if its supporting hyperplanes are in general position, meaning that each vertex is incident to $\dim(P)$ facets (or equivalently to $\dim(P)$ edges).

The (outer) \defn{normal cone} of a face~$F$ of~$P$ is the cone generated by the outer normal vectors of the facets of~$P$ containing~$F$.
In other words, it is the cone of vectors~$\b{c}$ such that the linear form~${\b{x} \mapsto \dotprod{\b{c}}{\b{x}}}$ on~$P$ is maximized by all points of the face~$F$.
The (outer) \defn{normal fan} of~$P$ is the collection of the (outer) normal cones of all its faces.
We say that a complete polyhedral fan~$\Fan$ in~$\R^d$ is \defn{polytopal} when it is the normal fan of a polytope~$P$ of~$\R^d$, and that~$P$ is a \defn{polytopal realization} of~$\Fan$.

@@

Note that~$\b{K}$ is only defined up to linear transformation, as it depends on the choice of a basis of the left kernel of~$\b{G}$. Note also that, by construction, the coefficients of the normal vectors of the facets of~$\typeCone(\Fan)$ arise from a linear dependence among the $\b{g}$-vectors, and hence belong to this kernel. When~$\typeCone(\Fan)$ is simplicial, its $(N-n)$ facets are necessarily linearly independent, and hence form a basis. They provide a particularly interesting choice of~$\b{K}$.

@@

\begin{figure}[htpb]
	\capstart
	\centerline{\includegraphics[scale=.8]{typeConeCoarsenings}}
	\caption{The type cone~$\ctypeCone(\Fan)$ of the fan~$\Fan$ of \cref{fig:typeCone}\,(left). On top of each face of $\ctypeCone(\Fan)$, we depict an example of polytope with the corresponding normal fan.}
	\label{fig:typeConeCoarsenings}
\end{figure}
\end{example}

@@

We now discuss the connection between convex combinations in the type cone and Minkowski sums of polytopes.
Recall that the \defn{Minkowski sum} of two polytopes~$P$ and~$Q$ is the polytope $P + Q \eqdef \set{p + q}{p \in P\text{ and } q \in Q}$.

@@

% \arnau{Can you please check the coordinates? C.\ Stump thought that two facets had some typo (a $\pm 1$ was in the wrong column)}

>>>>

% Type cone
\newcommand{\typeCone}{\mathbb{TC}} % type cone
\newcommand{\ctypeCone}{\smash{\overline{\mathbb{TC}}}} % type cone
\newcommand{\compatibilityDegree}[2]{(#1\,\|\,#2)} % compatibility degree
\newcommandx{\coefficient}[3][1={\ray[s]}, 2=\ray, 3=\ray']{\alpha_{#2,#3}(#1)} % coefficient in linear dependence

@@

Moreover, we use a similar approach to study the space of polytopal realizations of the \mbox{$\b{g}$-vector} fans of another generalization of the associahedron: non-kissing complexes (a.k.a. support $\tau$-tilting complexes) of gentle algebras. We show that the space of realizations of the non-kissing fan is simplicial when the gentle bound quiver is brick and $2$-acyclic, and we describe in this case its facet-defining inequalities in terms of mesh mutations.

@@

In the present paper, we explore specifically \textbf{gentle associahedra}. These complexes where constructed in~\cite{PaluPilaudPlamondon-nonkissing} as polytopal realizations of finite non-kissing complexes of gentle bound quivers. They encompass two families of simplicial complexes studied independently in the literature: on the one hand the grid associahedra introduced by T.~K.~Petersen, P.~Pylyavskyy and D.~Speyer in~\cite{PetersenPylyavskyySpeyer} for a staircase shape, studied by F.~Santos, C.~Stump and V.~Welker~\cite{SantosStumpWelker} for rectangular shapes, and extended by T.~McConville in~\cite{McConville} for arbitrary grid shapes; and on the other hand the Stokes polytopes and accordion associahedra studied by Y.~Baryshnikov~\cite{Baryshnikov}, F.~Chapoton~\cite{Chapoton-quadrangulations}, A.~Garver and T.~McConville~\cite{GarverMcConville} and T.~Manneville and V.~Pilaud~\cite{MannevillePilaud-accordion}. The latter, sometimes called accordiohedra, have also~recently appeared in the mathematical physics litterature concerning scattering amplitudes~\cite{BanerjeeLaddhaRaman-StokesPolytopes, Raman-PositiveGeometry}, with realizations inspired by those in~\cite{ArkaniHamedBaiHeYan}. It was shown in~\cite{PaluPilaudPlamondon-nonkissing, BrustleDouvilleMousavandThomasYildirim} that the non-kissing complex of a gentle bound quiver provides a combinatorial model for the support $\tau$-tilting complex~\cite{AdachiIyamaReiten} of the associated gentle algebra. Even for finite non-kissing complexes, it turns out that the type cone of the non-kissing fan is not always simplicial (see \cref{exm:nonSimplicialTypeConeNKC}). We prove however that the type cone is always simplicial when the gentle bound quiver is brick (at least $2$ relations in any cycle) and $2$-acyclic (no oriented cycle of length~$2$). Moreover, we precisely describe the facet-defining inequalities of the type cone in terms of linear dependences of $\b{g}$-vectors involved in certain mesh relations. We thus automatically derive a construction similar to that of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} describing all polytopal realizations of the $\b{g}$-vector fans of these brick and $2$-acyclic gentle bound quivers as sections of a positive orthant by well-chosen affine spaces (see \cref{subsec:typeConeCA}). Again, this result mainly relies on finding which relations minimally generate the relations between the $\b{g}$-vectors in the non-kissing complex. We present both a purely combinatorial proof (see \cref{subsec:typeConeCA}) and a purely algebraic proof for (non-necessarily gentle) brick algebras making use of extriangulated categories (see \cref{sec:extricats}).

@@

\smallskip
Further families of generalizations of the associahedron certainly deserve to be studied with our type cone approach.
We are investigating in particular graph associahedra~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}, brick polytopes~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and quotientopes~\cite{Reading-latticeCongruences, Reading-CambrianLattices, PilaudSantos-quotientopes}.
In fact, as already mentioned, it seems sometimes easier to find all polytopal realizations of a fan (by describing its type cone in terms of certain specific pairs of adjacent maximal cones) than to identify one specific explicit realization.
This is clearly the case when the type cone is simplicial as illustrated by the results of this paper: we naturally obtained all polytopal realizations of the $\b{g}$-vector fans of cluster algebras from cyclic seeds and of gentle algebras whose polytopality was only established very recently~\cite{HohlwegPilaudStella, PaluPilaudPlamondon-nonkissing}.
We believe that this approach gives reasonable hope that the polytopality of further fans could in the future be established using our type cone approach.
Tempting candidates are quotientopes for hyperplane arrangements beyond type~$A$, which are particularly interesting since the coarsenings of a fan essentially correspond to the faces of its type cone (see \cref{subsec:facesTypeCone}).

@@

\newtheorem*{coro:simplicialTypeCone}{\cref{coro:simplicialTypeCone}}
\begin{coro:simplicialTypeCone}
Assume that the type cone~$\typeCone(\Fan)$ is simplicial and let~$\b{K}$ be the $(N-n) \times N$-matrix whose rows are the inner normal vectors of the facets of~$\typeCone(\Fan)$. Then the polytope
\[
R_\b{\ell} \eqdef \set{\b{z} \in \R^N}{K\b{z} = \b{\ell} \text{ and } \b{z} \ge 0}
\]
is a realization of the fan~$\Fan$ for any positive vector~$\b{\ell} \in \R_{>0}^{N-n}$.
Moreover, the polytopes~$R_\b{\ell}$ for~$\b{\ell} \in \R_{>0}^{N-n}$ describe all polytopal realizations of~$\Fan$.
\end{coro:simplicialTypeCone}

@@

\newpage
\para{Schematic diagram of logical dependencies}

@@

\vspace{.5cm}
\tikzset{
  double arrow/.style={
    -latex, line width=3pt, black, % first arrow
    postaction={draw,line width=2pt, white, shorten <= 1,shorten >= 2}, % second arrow
  }
}
\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
	\node[diamond,aspect=3, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=-2] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center, minimum width = 5.3cm, minimum height=1.2cm] (I21) at (-6,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5.3cm, minimum width = 5cm, minimum height=1.2cm] (I22) at (0,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5.3cm, minimum width = 5cm, minimum height=1.2cm] (I23) at (6,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=5cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=5cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(-1.6,-1.5)}]\textit{applied to}}]{} (I21.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=3}, draw=blue, fill=blue!10] (I1.270) to node[label={[shift={(1,-1.)}]\textit{applied to}}]{} (I22.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.360) to [bend left, looseness=1.1] node[label={[shift={(1.6,-1.5)}]\textit{applied to}}]{} (I23.90);
	\draw[thick, double distance=2pt, -implies, in=270] (I22.270)+(-1.5,-1.2) to node[semithick, right, outer sep=0, pos=0.55]{\textit{used for}} ([yshift=.2cm] I22.270);
	\draw[thick, double distance=2pt, -implies, in=270] (I23.270)+(-1.5,-1.2) to node[semithick, right, outer sep=0, pos=0.55]{\textit{used for}} ([yshift=.2cm] I23.270);
	\draw[thick, ->] (II1.0) to node[semithick, below, outer sep=0]{\textit{generalizes to}} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=([shift={(0,-.15)}]I22.270), \p{right}=(I23.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I23.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}

@@

A hyperplane~$H \subset \R^n$ is a \defn{supporting hyperplane} of a set~$X \subset \R^n$ if~$H \cap X \ne \varnothing$ and~$X$ is contained in one of the two closed half-spaces of~$\R^n$ defined by~$H$.

We denote by~$\R_{\ge0}\rays \eqdef \set{\sum_{\ray \in \rays} \lambda_{\ray} \, \ray}{\lambda_{\ray} \in \R_{\ge0}}$ the \defn{positive span} of a set~$\rays$ of vectors of~$\R^n$.
A \defn{polyhedral cone} is a subset of~$\R^n$ defined equivalently as the positive span of finitely many vectors or as the intersection of finitely many closed linear halfspaces.
The \defn{faces} of a cone~$C$ are the intersections of~$C$ with the supporting hyperplanes of~$C$.
The $1$-dimensional (resp.~codimension~$1$) faces of~$C$ are called~\defn{rays} (resp.~\defn{facets}) of~$C$.
A cone is \defn{simplicial} if it is generated by a set of linearly independent vectors.

A \defn{polyhedral fan} is a collection~$\Fan$ of polyhedral cones such that
\begin{itemize}
\item if~$C \in \Fan$ and~$F$ is a face of~$C$, then~$F \in \Fan$,
\item the intersection of any two cones of~$\Fan$ is a face of both.
\end{itemize}
A fan is \defn{simplicial} if all its cones are, \defn{complete} if the union of its cones covers the ambient space~$\R^n$, and \defn{essential} if it contains the cone~$\{\b{0}\}$. Note that every complete fan is the product of an essential fan with its lineality space.
For two fans~$\Fan, \cal{G}$ in~$\R^n$, we say that~$\Fan$ \defn{refines}~$\cal{G}$ (and that~$\cal{G}$ \defn{coarsens}~$\Fan$) if every cone of~$\Fan$ is contained in a cone of~$\cal{G}$.

A \defn{polytope} is a subset~$P$ of~$\R^n$ defined equivalently as the convex hull of finitely many points or as a bounded intersection of finitely many closed affine halfspaces.
The \defn{dimension}~$\dim(P)$ is the dimension of the affine hull of~$P$.
The \defn{faces} of~$P$ are the intersections of~$P$ with its supporting hyperplanes.
The dimension~$0$ (resp.~dimension~$1$, resp.~codimension~$1$) faces are called \defn{vertices} (resp.~\defn{edges}, resp.~\defn{facets}) of~$P$.
A polytope is \defn{simple} if its supporting hyperplanes are in general position, meaning that each vertex is incident to $\dim(P)$ facets (or equivalently to $\dim(P)$ edges).

The (outer) \defn{normal cone} of a face~$F$ of~$P$ is the cone generated by the outer normal vectors of the facets of~$P$ containing~$F$.
In other words, it is the cone of vectors~$\b{c}$ such that the linear form~${\b{x} \mapsto \dotprod{\b{c}}{\b{x}}}$ on~$P$ is maximized by all points of the face~$F$.
The (outer) \defn{normal fan} of~$P$ is the collection of the (outer) normal cones of all its faces.
We say that a complete polyhedral fan~$\Fan$ in~$\R^n$ is \defn{polytopal} when it is the normal fan of a polytope~$P$ of~$\R^n$, and that~$P$ is a \defn{polytopal realization} of~$\Fan$.

@@

Note that~$\b{K}$ is only defined up to linear transformation, as it depends on the choice of a basis of the left kernel of~$\b{G}$. Note also that, by construction, the coefficients of the normal vectors of the facets of~$\typeCone(\Fan)$ arise from a linear dependence among the rays of~$\Fan$, and hence belong to this kernel. When~$\typeCone(\Fan)$ is simplicial, its $(N-n)$ facets are necessarily linearly independent, and hence form a basis. They provide a particularly interesting choice of~$\b{K}$.

@@

\begin{figure}[htpb]
	\capstart
	\centerline{\includegraphics[scale=.7]{typeConeCoarsenings}}
	\caption{The type cone~$\ctypeCone(\Fan)$ of the fan~$\Fan$ of \cref{fig:typeCone}\,(left). On top of each face of $\ctypeCone(\Fan)$, we depict an example of polytope with the corresponding normal fan.}
	\label{fig:typeConeCoarsenings}
\end{figure}
\end{example}

@@

\enlargethispage{.5cm}
We now discuss the connection between convex combinations in the type cone and Minkowski sums of polytopes.
Recall that the \defn{Minkowski sum} of two polytopes~$P$ and~$Q$ is the polytope $P + Q \eqdef \set{p + q}{p \in P\text{ and } q \in Q}$.

@@
